---
output:
  html_document: default

title: 'Tourism Model Inputs'
author: "Kat Millage"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
---

```{r setup, include=FALSE}
# Source common R file
source(here::here("common.R"))

# Do we need to rematch our ocean pixel raster to EEZs? This is time consuming, so if the pixel resolution hasn't changed it's best to skip this. 
rematch_eez <- F
```

# Introduction

This script creates the supply and demand inputs for the tourism model:
- Dive suitability by cell
- Number of extrapolated dives by cell
- Price by cell (global median price)

It also generates a couple of additional inputs for use in the sensitivity analysis:
- Price by cell (median price by country/region)
- Price by cell (spatially interpolated median price)

# Load and Process Data

```{r}
# Let's turn our ocean data frame back into a raster stack...
ocean_low_res_moll <- raster::raster(file.path(emlab_data_dir, "ocean-conservation-priorities", "inputs", "ocean_low_res_moll.tif"))

# Get our cell ids
ocean_df <- ocean_low_res_moll %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("lon", "lat", "ocean")) %>% 
  filter(!is.na(ocean)) %>% 
  rowid_to_column(var = "cell_id") %>% 
  as_tibble() %>%
  relocate(lon, lat, cell_id, ocean)

write_csv(ocean_df, file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "ocean_moll_df.csv"))

# Now we need to turn this back into a raster brick
ocean_raster <- rasterFromXYZ(ocean_df)

plot(ocean_raster, "ocean")
```

```{r}
# Get our non-ocean cells and give them an id too...
land_df <- ocean_low_res_moll %>%
  raster::as.data.frame(xy = T) %>%
  set_names(c("lon", "lat", "ocean")) %>%
  filter(is.na(ocean))

land_df$cell_id <- seq(max(ocean_df$cell_id)+1, max(ocean_df$cell_id)+nrow(land_df), by = 1)

land_df <- land_df %>%
  mutate(ocean = 0) %>%
  as_tibble() %>%
  relocate(lon, lat, cell_id, ocean)

# Now we need to turn this back into a raster brick
land_raster <- rasterFromXYZ(land_df)

plot(land_raster, "ocean")

# Now let's lump them back together...
world_df <- ocean_df %>%
  bind_rows(land_df) %>%
  as_tibble() %>%
  relocate(lon, lat, cell_id, ocean)

write_csv(world_df, file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "world_moll_df.csv"))

# Now we need to turn this back into a raster brick
world_raster <- rasterFromXYZ(world_df)

plot(world_raster, "ocean")
```

```{r}
# Load the cleaned dive database
dive_dat <- read_csv(file.path(emlab_project_dir, "data", "02-processed", "data-prep", "diveboard_database_extrapolated_2010_2020_final.csv"))

# Load cleaned price data
price_dat <- read_csv(file.path(emlab_project_dir, "data", "02-processed", "data-prep", "prices_day_trips_clean.csv"))
```

```{r}
# Extract all of our dives cooresponding to marine sites
dive_dat_marine <- dive_dat %>%
  dplyr::filter(site_classification == "Marine")
```

# Inputs #1-2: Dive Suitability and Number of Dives

```{r}
# Rename
dives_sites_out <- dive_dat_marine 

# Now let's convert the coordinate system on our dive spot data so it matches up with our ocean pixels... 
dives_spdf <- SpatialPointsDataFrame(cbind(dives_sites_out$lon, dives_sites_out$lat), as.data.frame(dives_sites_out$dive_id), proj4string=CRS("+proj=longlat"))

dives_spdf_moll <- spTransform(dives_spdf, CRS = prj_moll)

# Now we want to overlay the ocean raster onto our points so we get the ocean cell corresponding to each dive  
dives_matched_to_cell_id <- cbind(dives_spdf_moll, raster::extract(ocean_raster, dives_spdf_moll)) %>%
  as.data.frame() %>%
  rename(dive_id = `dives_sites_out.dive_id`,
         x_coord = `coords.x1`,
         y_coord = `coords.x2`)

# I think it would be useful to add info we want back in here... 
dives_matched_out <- dives_matched_to_cell_id %>%
  left_join(dives_sites_out, by = "dive_id")

write_csv(dives_matched_out, file.path(emlab_project_dir, "data", "diagnostics", "marine_dives_matched_to_cell_id_raw.csv"))
```

Figure out which of our marine dives didn't match to a cell and find the nearest pixel. Since the model pixels don't cover every inch of EEZs, we're going to allocate those dives to their nearest ocean pixel so they aren't excluded from the analysis. 

```{r}
# Find marine dives with no matched cell
marine_dives_no_cell <- dives_matched_out %>%
  dplyr::filter(is.na(cell_id))

# Convert to spatial object
dives_no_cell_spdf <- SpatialPointsDataFrame(cbind(marine_dives_no_cell$lon, marine_dives_no_cell$lat), as.data.frame(marine_dives_no_cell$dive_id), proj4string=CRS("+proj=longlat"))

dives_no_cell_spdf_moll<- spTransform(dives_no_cell_spdf, CRS = prj_moll)

# Output df
dives_no_cell_match <- tibble()

for(i in 1:nrow(marine_dives_no_cell)){
  # Get distances
  dist <- distanceFromPoints(ocean_raster, dives_no_cell_spdf_moll[i,])
  
  # Match back to cell ids and find the ocean cell with the shortest distance
  dist_df <- dist %>%
    raster::as.data.frame(xy = T) %>%
    set_names(c("lon", "lat", "distance")) %>%
    left_join(ocean_df, by = c("lon", "lat")) %>%
    dplyr::filter(!is.na(cell_id) & ocean == 1) %>%
    dplyr::filter(distance == min(distance, na.rm = T))

  o <- cbind("dive_id" = marine_dives_no_cell$dive_id[1], dist_df)
  # Add 
  dives_no_cell_match <- dives_no_cell_match %>%
    bind_rows(o)
}

# Save
dives_no_cell_match_out <- dives_no_cell_match %>%
  dplyr::select(dive_id, cell_id, ocean, x_coord = lon, y_coord = lat)

# Now let's remove those dives from our original df and add the newly matched cells back in
dives_matched_out_fixed <- dives_matched_to_cell_id %>%
  dplyr::filter(!dive_id %in% marine_dives_no_cell$dive_id) %>%
  bind_rows(dives_no_cell_match_out) %>%
  left_join(dives_sites_out, by = "dive_id")

# Save
write_csv(dives_matched_out_fixed, file.path(emlab_project_dir, "data", "02-processed", "data-prep", "marine_dives_matched_to_cell_id_all.csv"))
```

## Aggregate by Cell and Make Inputs

Now time to aggregate at the cell level and make our rasters. First let's summarize our raw number of dives data by cell...

```{r}
# Let's double check that we don't have any not falling within our ocean pixels. We're good. 
dives_matched_out_marine <- dives_matched_out_fixed %>%
  dplyr::filter(!is.na(ocean) & ocean == 1 & !is.na(cell_id)) %>%
  ungroup()

write_csv(dives_matched_out_marine, file.path(emlab_project_dir, "data", "02-processed", "data-prep", "marine_dives_matched_to_cell_id_ocean_only.csv"))

# Just our of curiosity, let's see what we have by cell and year
dives_by_cell_year_marine <- dives_matched_out_marine %>%
  group_by(year, cell_id) %>%
  summarize(n_spots = n_distinct(spot_id),
            n_dives = n_distinct(dive_id),
            n_dives_extrap = sum(n_dives_extrap, na.rm = T),
            n_dives_extrap_min = sum(n_dives_extrap_min, na.rm = T),
            n_dives_extrap_max = sum(n_dives_extrap_max, na.rm = T)) %>%
  ungroup() %>%
  arrange(year, cell_id)

write_csv(dives_by_cell_year_marine, 
          file.path(emlab_project_dir, "data", "diagnostics", "number_of_marine_sites_and_dives_by_cell_and_year_raw.csv"))
```

Now let's calculate the total number of marine dives we ended up with and the range

```{r}
# Global estimate of the number of dives
n_global_dives <- sum(dive_dat$n_dives_extrap, na.rm = T)
n_global_dives_range <- c(sum(dive_dat$n_dives_extrap_min, na.rm = T),
                          sum(dive_dat$n_dives_extrap_max, na.rm = T))

# See what we ended up with in our raster
n_dives_model <- sum(dives_by_cell_year_marine$n_dives_extrap, na.rm = T) # 33,079,176
n_dives_model_range <- c(sum(dives_by_cell_year_marine$n_dives_extrap_min, na.rm = T), # 17,100,252
                         sum(dives_by_cell_year_marine$n_dives_extrap_max, na.rm = T)) # 53,963,910

# Proportion marine
prop_marine <- n_dives_model/n_global_dives # 65.26%
```

### Number of dives

```{r}
# Now let's sum across all years and get the dives per cell
dives_by_cell_marine <- dives_by_cell_year_marine %>%
  group_by(cell_id) %>%
  summarize(n_spots = sum(n_spots, na.rm = T),
            n_dives = sum(n_dives, na.rm = T),
            n_dives_extrap = sum(n_dives_extrap, na.rm = T),
            n_dives_extrap_min = sum(n_dives_extrap_min, na.rm = T),
            n_dives_extrap_max = sum(n_dives_extrap_max, na.rm = T)) %>%
  ungroup() %>%
  arrange(cell_id)

# Make sure we still have the same total
n_dives_model_diff <- n_dives_model - sum(dives_by_cell_marine$n_dives_extrap, na.rm = T) 

# And finally, match back to our full ocean_df 
dives_by_cell_out <- ocean_df %>%
  left_join(dives_by_cell_marine, by = "cell_id")

# Save
write_csv(dives_by_cell_out %>% dplyr::select(-lat, -lon, -ocean), file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "number_of_marine_sites_and_dives_by_cell.csv"))
```

```{r}
barwidth = 15
barheight = 1

dive_extrap_breaks = c(100,1000,1e4,1e5,1e6)
```

```{r}
# Plot
dives_by_cell_input_plot <- dives_by_cell_out %>%
  dplyr::filter(!is.na(n_dives_extrap)) %>%
  ggplot()+
  map_layer_moll+
  geom_raster(aes(x = lon, y = lat, fill = n_dives_extrap))+
  scale_fill_viridis(name = "Dives (thousands, extrapolated)",
                      trans = "log",
                      breaks = dive_extrap_breaks, 
                      labels = dive_extrap_breaks/1e4,
                      limits = c(100,1e6),
                     option = "D",
                     guide = guide_colorbar(title.position = "top", title.hjust = 0.5, barwidth = barwidth, barheight = barheight))+
  theme_basic_map()+
  labs(x = "", y = "", title = paste0("Total number of marine dives included in model pixels (extrapolated): ",
                                      format(sum(dives_by_cell_out$n_dives_extrap, na.rm = T), big.mark = ",")))
  
# Output
save_plots(dives_by_cell_input_plot, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "number_of_dives_by_cell.png"),
           width_in = 7.5,
           height_in = 4.5)

dives_by_cell_input_plot
```

### Dive suitability

Make a dive site suitability input now... 

```{r}
# Now summarize by cell across all years
spots_by_cell <- dives_by_cell_out %>%
  mutate(suitable = ifelse(!is.na(n_spots) & n_spots >= 1, 1, 0))

write_csv(spots_by_cell %>% dplyr::select(cell_id, suitable), file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "dive_suitability_by_cell.csv"))
```

```{r}
# Let's get the ocean area that is suitable
total_ocean_area <- n_distinct(ocean_df$cell_id[ocean_df$ocean == 1])*2500
suitable_area <- n_distinct(spots_by_cell$cell_id[spots_by_cell$suitable == 1])*2500

# Plot
suitability_by_cell_input_plot <- spots_by_cell %>%
  mutate(suitability = case_when(suitable == 0 ~ "No",
                              suitable == 1 ~ "Yes")) %>%
  ggplot()+
  map_layer_moll+
  geom_raster(aes(x = lon, y = lat, fill = suitability))+
  guides(fill = guide_legend(title = "Suitable for Recreational Diving?"))+
  theme_basic_map()+
  scale_fill_manual(values = c(NA, "aquamarine3"), na.value = NA)+
  labs(x = "", y = "",
       title = paste0("Total ocean area suitable for diving: ", percent(suitable_area/total_ocean_area, accuracy = 0.01)))+
  theme(legend.key=element_rect(colour="black"))
  
# Output
save_plots(suitability_by_cell_input_plot, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "dive_suitability_by_cell.png"),
           width_in = 7.5,
           height_in = 4.5)

suitability_by_cell_input_plot
```

```{r}
# Turn into a real raster so we can use it for masking
spots_raster <- spots_by_cell %>%
  dplyr::select(lon, lat, suitable) %>%
  rasterFromXYZ()

plot(spots_raster, "suitable")
writeRaster(spots_raster, file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "dive_suitability_by_cell_raster.tif"), overwrite = T)
```

# Input #3: Price

## Match to cells

```{r}
# Assign row identifier to the price data so we can match up later... 
price_df <- price_dat %>%
  rename(shop_id = ID) %>%
  rowid_to_column(var = "row_id")

# Now let's convert the coordinate system on our price data so it matches up with our ocean pixels... 
price_df_filtered <- price_df %>% dplyr::select(lon=longitude, lat=latitude, row_id) %>%
  dplyr::filter(lon >= -180 & lon <= 180) %>% 
  dplyr::filter(lat >= -90 & lat <=90)

prices_spdf <- SpatialPointsDataFrame(cbind(price_df_filtered$lon, price_df_filtered$lat), as.data.frame(price_df_filtered$row_id), proj4string=CRS("+proj=longlat"))

prices_spdf_moll <- spTransform(prices_spdf, CRS = prj_moll)

# Now we want to try to overlay the global raster onto our points...  
prices_matched_to_cell_id <- cbind(prices_spdf_moll, raster::extract(world_raster, prices_spdf_moll)) %>%
  as.data.frame() %>%
  rename(row_id = `price_df_filtered.row_id`,
         lon = `coords.x1`,
         lat = `coords.x2`)

# Let's add our complete price data back in before saving... 
prices_matched_out <- prices_matched_to_cell_id %>%
  left_join(price_df %>% dplyr::select(-longitude, -latitude), by = "row_id")

write_csv(prices_matched_out, file.path(emlab_project_dir, "data", "02-processed", "data-prep", "prices_matched_to_cell_id_world.csv"))
```

## Calculate Aggregate Prices and Make Inputs

Since we often have multiple prices from the same shop, let's get our "best" per dive price and then aggregate only those. 

```{r}
# Let's first get the entries where we have 1-tank data - 262 entries
price_dat_1_dive <- prices_matched_out %>%
  dplyr::filter(description %in% c("1 Dive")) %>%
  mutate(per_dive_price_best = price_converted)

# Then 2-tank data for operators where we don't have 1-tank data - 102 entries
price_dat_2_dive <- prices_matched_out %>%
  dplyr::filter(description %in% c("2 Dives")) %>%
  dplyr::filter(!(shop_id %in% price_dat_1_dive$shop_id)) %>%
  mutate(per_dive_price_best = price_converted/2)

# Then N-tank data for operators where we don't have 1-tank or 2-tank data - 6 entries
price_dat_n_dive <- prices_matched_out %>%
  dplyr::filter(description %in% c("N Dives")) %>%
  dplyr::filter(!(shop_id %in% price_dat_1_dive$shop_id) & !(shop_id %in% price_dat_2_dive$shop_id)) %>%
  mutate(per_dive_price_best = price_converted/n_dives) %>%
  dplyr::filter(!is.na(per_dive_price_best))

# Combine - 370 entries total
price_dat_per_dive <- price_dat_1_dive %>%
  bind_rows(price_dat_2_dive) %>%
  bind_rows(price_dat_n_dive)
```

### Globally Constant Price

```{r}
# Global median price - $58.75
median_single_price_globally <- price_dat_per_dive %>%
  dplyr::filter(is_liveaboard == F) %>% # just in case
  summarize(median_single_price_global = median(per_dive_price_best, na.rm = T)) %>%
  dplyr::pull(median_single_price_global)

## Make raster
price_constant <- spots_by_cell %>%
  mutate(price = ifelse(suitable == 1, median_single_price_globally, NA)) %>%
  dplyr::select(cell_id, price)

write_csv(price_constant, file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "price_constant_by_cell.csv"))
```

```{r}
price_breaks <- c(0,30,60,90,120)
price_limits <- c(0,120)
```

```{r}
# Now we match back to our ocean df and plot
prices_constant_map <- price_constant %>%
  right_join(ocean_df, by = "cell_id") %>%
  dplyr::filter(!is.na(price)) %>%
  ggplot()+
  map_layer_moll+
  geom_raster(aes(x = lon, y = lat, fill = price))+
  scale_fill_viridis(name = "Price per Dive (USD)",
                      breaks = price_breaks,
                      limits = price_limits,
                     option = "C",
                     guide = guide_colorbar(title.position = "top", title.hjust = 0.5, barwidth = barwidth, barheight = barheight))+
  theme_basic_map()+
  labs(x = "", y = "", title = paste0("Median dive price for all model pixels (USD): ",
                                      dollar(median(price_constant$price, na.rm = T))))
  
# Output
save_plots(prices_constant_map, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "price_constant_by_cell.png"),
           width_in = 7.5,
           height_in = 4.5)

prices_constant_map
```

### Constant Within each Country (or Region)

The second price input we're going to make is one for the sensitivty analysis. Instead of assuming a constant global price, we're going to apply median prices by country or by region (if we don't have price data for that country) to all pixels under that country's jurisdiction. 

```{r}
# By region
median_single_price_region <- price_dat_per_dive %>%
  dplyr::filter(is_liveaboard == F) %>% # just in case
  mutate(region = countrycode(country_name, "country.name", "region")) %>%
  group_by(region) %>%
  summarize(median_single_price_continent = median(per_dive_price_best, na.rm = T)) %>%
  ungroup()

# By country
median_single_price_country <- price_dat_per_dive %>%
  dplyr::filter(is_liveaboard == F) %>% # just in case
  mutate(country_code_a3 = countrycode(country_name, "country.name", "iso3c")) %>%
  group_by(country_code_a3) %>%
  summarize(median_single_price_country = median(per_dive_price_best, na.rm = T)) %>%
  ungroup()
```

Approach #1: We need to overlay the EEZ boundaries over our ocean pixels to figure out the best administering country for each ocean pixel. We're going to do this based on the center coordinates of each cell. This process is pretty time consuming (3+ hours), so it is suggested to not rerun this portion unless absolutely needed. This shouldn't change unless a different version of the EEZ boundaries needs to be used or our pixel resolution changes. 


```{r}
shp2raster <- function(shp, mask.raster, label, value, transform = FALSE, proj.from = NA,
    proj.to = NA, map = TRUE) {
    require(raster, rgdal)
    # use transform==TRUE if the polygon is not in the same coordinate system as
    # the output raster, setting proj.from & proj.to to the appropriate
    # projections
    if (transform == TRUE) {
        proj4string(shp) <- proj.from
        shp <- spTransform(shp, proj.to)
    }
    # convert the shapefile to a raster based on a standardized background raster
    r <- rasterize(shp, mask.raster)
                   
    # set the cells associated with the shapefile to the specified value
    r[!is.na(r)] <- value
    # merge the new raster with the mask raster and export to the working
    # directory as a tif file
    r <- mask(merge(r, mask.raster), mask.raster, filename = label, format = "GTiff",
        overwrite = T)
    # plot map of new raster
    if (map == TRUE) {
        plot(r, main = label, axes = F, box = F)
    }
    names(r) <- label
    return(r)
}
```

```{r}
if(rematch_eez){
  
# Get EEZ data (marine regions v11)
eez_sf <- st_read(file.path(emlab_data_dir, "marine-regions-eez-v11", "World_EEZ_v11_20191118_gpkg", "eez_v11.gpkg")) %>%
  janitor::clean_names() %>%
  dplyr::select(mrgid,
                eez_name = geoname,
                eez_type = pol_type,
                eez_ter1_name = territory1,
                eez_ter1_iso3 = iso_ter1,
                eez_sov1_name = sovereign1,
                eez_ter2_name = territory2,
                eez_ter2_iso3 = iso_ter2,
                eez_sov2_name = sovereign2,
                eez_ter3_name = territory3,
                eez_ter3_iso3 = iso_ter3,
                eez_sov3_name = sovereign3,
                eez_area_km = area_km2) %>%
  sf::st_transform(crs = st_crs(prj_moll))

# Fix invalid geometries
eez_sf <- st_make_valid(eez_sf)

# Figure out which are still not valid - none, we're good
which(!st_is_valid(eez_sf))
  
# Convert the EEZ shapefile to a raster based on the suitability raster
r <- rasterize(eez_sf, spots_raster)

plot(r, "mrgid")

r1 <- r %>%
  as.data.frame(xy=T)

# Get EEZ code for each lat/lon
r_eez <- r1 %>%
  dplyr::filter(!is.na(layer_mrgid)) %>%
  group_by(x,y) %>%
  summarize(mrgid = unique(layer_mrgid)) %>%
  rename(lon = x, lat = y)

# Now try matching back to our dives by cell and see if we end up with any cells with dives but no EEZ
test <- dives_by_cell_out %>%
  left_join(r_eez, by = c("lon", "lat")) # Same number of rows

# Do we have any cells with dives but no EEZ? Yes, 399. 
test2 <- test %>%
  dplyr::filter(!is.na(n_dives_extrap) & is.na(mrgid))

}
```


```{r}
if(rematch_eez){
  
# Get EEZ data (marine regions v11)
eez_sf <- st_read(file.path(emlab_data_dir, "marine-regions-eez-v11", "World_EEZ_v11_20191118_gpkg", "eez_v11.gpkg")) %>%
  janitor::clean_names() %>%
  dplyr::select(mrgid,
                eez_name = geoname,
                eez_type = pol_type,
                eez_ter1_name = territory1,
                eez_ter1_iso3 = iso_ter1,
                eez_sov1_name = sovereign1,
                eez_ter2_name = territory2,
                eez_ter2_iso3 = iso_ter2,
                eez_sov2_name = sovereign2,
                eez_ter3_name = territory3,
                eez_ter3_iso3 = iso_ter3,
                eez_sov3_name = sovereign3,
                eez_area_km = area_km2) %>%
  sf::st_transform(crs = st_crs(prj_moll))

# Fix invalid geometries
eez_sf <- st_make_valid(eez_sf)

# Figure out which are still not valid
which(!st_is_valid(eez_sf))

# Get cell coordinates
cell_coords <- ocean_df %>%
  distinct(cell_id, lat, lon)

# Make coordinates into spatial points
cell_coords_sf <- cell_coords %>%
  st_as_sf(coords = c("lon", "lat"),
           remove = T)

# Convert coords
st_crs(cell_coords_sf) <- st_crs(eez_sf)

# Find cells within EEZ boundaries
cells_by_eez <- st_join(cell_coords_sf, eez_sf, join = st_within)

# Match back to all pixels (HS are NA)
cells_by_eez_all <- cells_by_eez %>%
  st_drop_geometry() %>%
  right_join(ocean_df, by = "cell_id")

}
```

We apparently need to deal with a few duplicates because of overlapping areas. Examples include: 
- Jan Mayen Exclusive Economic Zone (8437) and Joint regime area Iceland / Norway (Jan Mayen) (48975)
- Icelandic Exclusive Economic Zone (5680) and Joint regime area Iceland / Norway (Jan Mayen) (48975)
- United Kingdom Exclusive Economic Zone (5696) and Joint regime area United Kingdom / Denmark (Faeroe Islands) (48967)

There are more, but it looks like we can just give priority to the 200NM area in all cases. 

```{r}
if(rematch_eez){
  
# Find duplicate cells
cell_duplicates <- cells_by_eez_all %>%
  dplyr::filter(!is.na(mrgid)) %>%
  group_by(cell_id) %>%
  count() %>%
  dplyr::filter(n > 1)

# Get details - we need to get this down to 88
cell_duplicate_details <- cells_by_eez_all %>%
  dplyr::filter(cell_id %in% c(cell_duplicates$cell_id)) %>%
  group_by(cell_id) %>%
  mutate(choose = case_when(eez_type == "200NM" ~ "Y",
                            eez_type == "Joint regime" ~ "N")) %>%
  ungroup() %>%
  dplyr::filter(choose == "Y")

# Now join back together
cells_by_eez_out <- cells_by_eez_all %>%
  dplyr::filter(!(cell_id %in% cell_duplicate_details$cell_id)) %>%
  bind_rows(cell_duplicate_details) %>%
  arrange(cell_id) %>%
  mutate(region = countrycode(eez_ter1_iso3, "iso3c", "region")) %>%
  dplyr::select(-choose)

# Save
write_csv(cells_by_eez_out, file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "ocean_moll_df_with_eezs.csv"))

}else{
  
  cells_by_eez_out <- read_csv(file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "files", "ocean_moll_df_with_eezs.csv"))
  
}
```

Now to match. 

```{r}
# Join price data back
prices_by_cell_country_region <- spots_by_cell %>%
  dplyr::select(cell_id, suitable) %>%
  left_join(cells_by_eez_out %>% dplyr::select(cell_id, mrgid, eez_name, eez_ter1_iso3, region), by = "cell_id") %>%
  left_join(median_single_price_country, by = c("eez_ter1_iso3" = "country_code_a3")) %>%
  left_join(median_single_price_region, by = "region") %>%
  mutate(price = case_when(!is.na(mrgid) & !is.na(median_single_price_country) ~ median_single_price_country,
                           !is.na(mrgid) & !is.na(median_single_price_continent) ~ median_single_price_continent,
                           !is.na(mrgid) ~ median_single_price_globally,
                           suitable == 1 ~ median_single_price_globally)) %>%
  dplyr::select(cell_id, suitable, price, eez_ter1_iso3, region)

# Let's see if we have any missing values. We're mostly concerned about areas we've deemed suitable that don't have a price
price_test1 <- prices_by_cell_country_region %>% dplyr::filter(suitable == 1 & is.na(price))
# Looks like we're good. 
```

Let's map this before removing prices for non-suitable pixels.

```{r}
# Now we match back to our ocean df and plot
prices_country_region_map <- prices_by_cell_country_region %>%
  right_join(ocean_df, by = "cell_id") %>%
  dplyr::filter(!is.na(price)) %>%
  ggplot()+
  map_layer_moll+
  geom_raster(aes(x = lon, y = lat, fill = price))+
  scale_fill_viridis(name = "Price per Dive (USD)",
                      breaks = price_breaks,
                      limits = price_limits,
                     option = "C",
                     guide = guide_colorbar(title.position = "top", title.hjust = 0.5, barwidth = barwidth, barheight = barheight))+
  theme_basic_map()+
  labs(x = "", y = "", title = paste0("Median dive price for all EEZ pixels (USD): ",
                                      dollar(median(prices_by_cell_country_region$price, na.rm = T))))
  
# Output
save_plots(prices_country_region_map, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "price_country_region_all_by_cell.png"),
           width_in = 7.5,
           height_in = 4.5)

prices_country_region_map
```

Now, remove prices for non-suitable pixels, plot again, and save. 

```{r}
# Filter and save
prices_by_cell_country_region_out <- prices_by_cell_country_region %>%
  mutate(price = ifelse(suitable == 1, price, NA))

write_csv(prices_by_cell_country_region_out, file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "price_country_region_by_cell.csv"))

# Now we match back to our ocean df and plot
prices_country_region_suitable_map <- prices_by_cell_country_region_out %>%
  right_join(ocean_df, by = "cell_id") %>%
  dplyr::filter(!is.na(price)) %>%
  ggplot()+
  map_layer_moll+
  geom_raster(aes(x = lon, y = lat, fill = price))+
  scale_fill_viridis(name = "Price per Dive (USD)",
                      breaks = price_breaks,
                      limits = price_limits,
                     option = "C",
                     guide = guide_colorbar(title.position = "top", title.hjust = 0.5, barwidth = barwidth, barheight = barheight))+
  theme_basic_map()+
  labs(x = "", y = "", title = paste0("Median dive price for all model pixels (USD): ",
                                      dollar(median(prices_by_cell_country_region_out$price, na.rm = T))))
  
# Output
save_plots(prices_country_region_suitable_map, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "price_country_region_suitable_by_cell.png"),
           width_in = 7.5,
           height_in = 4.5)

prices_country_region_suitable_map
```

### Interpolated

Our last price input (also for the sensitivity analysis) involves spatially interpolating prices based on distance from the dive operators. 

```{r}
median_prices_by_world_cell <- price_dat_per_dive %>%
  group_by(cell_id) %>%
  summarize(median_price = median(per_dive_price_best, na.rm = T)) %>%
  ungroup()
```

```{r}
# Join back to raster
world_df_w_prices <- world_df %>%
  left_join(median_prices_by_world_cell, by = "cell_id")

price_raster_raw <- rasterFromXYZ(world_df_w_prices)
plot(price_raster_raw, "median_price")

# Extract just coords
coord <- world_df_w_prices %>% dplyr::select(lon, lat)
price <- world_df_w_prices$median_price

# Remove NA's
i <- !is.na(world_df_w_prices$median_price)
xy <- coord[i,]
v <- price[i]

# Thin plate spline model
library(fields) 
tps <- Tps(xy, v)
p <- raster(price_raster_raw)

# Use model to predict values at all locations
p <- interpolate(p, tps)
p <- mask(p, price_raster_raw, maskvalue = 0)

# Plot interpolated raster
plot(p)

# Convert back to DF and match to cell ids
price_interpolated_df <- p %>% 
  raster::as.data.frame(xy = T) %>% 
  set_names(c("lon", "lat", "price", "ocean", "unknown")) %>% 
  dplyr::select(lon, lat, price) %>%
  left_join(world_df, by = c("lon", "lat")) %>%
  as_tibble() %>%
  relocate(lon, lat, cell_id, ocean, price)

# Save raw interpolation
write_csv(price_interpolated_df, file.path(emlab_project_dir, "data", "diagnostics", "prices_interpolated_matched_to_world_cell_id.csv"))
```

```{r}
# Now we match back to our ocean df and plot
prices_interpolated_map_all <- price_interpolated_df %>%
  dplyr::select(-lon, -lat) %>%
  dplyr::filter(ocean != 0) %>%
  right_join(ocean_df, by = "cell_id") %>%
  dplyr::filter(!is.na(price)) %>%
  ggplot()+
  map_layer_moll+
  geom_raster(aes(x = lon, y = lat, fill = price))+
  scale_fill_viridis(name = "Price per Dive (USD)",
                      breaks = price_breaks,
                      limits = price_limits,
                     oob = squish,
                     option = "C",
                     guide = guide_colorbar(title.position = "top", title.hjust = 0.5, barwidth = barwidth, barheight = barheight))+
  theme_basic_map()+
  labs(x = "", y = "", title = paste0("Median dive price for all ocean pixels (USD): ",
                                      dollar(median(price_interpolated_df$price[price_interpolated_df$ocean ==1], na.rm = T))))
  
# Output
save_plots(prices_interpolated_map_all, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "price_interpolated_all_by_cell.png"),
           width_in = 7.5,
           height_in = 4.5)

prices_interpolated_map_all
```

Finally, we need to limit our interpolated prices to the cells deemed suitable. 

```{r}
# Limit prices to suitable cells... 
price_interpolated_out <- price_interpolated_df %>%
  dplyr::filter(ocean != 0) %>%
  dplyr::select(cell_id, price) %>%
  right_join(ocean_df, by = "cell_id") %>%
  left_join(spots_by_cell %>% dplyr::select(cell_id, suitable)) %>%
  mutate(price = ifelse(suitable == 1, price, NA)) %>%
  dplyr::select(cell_id, suitable, price)

write_csv(price_interpolated_out, file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "price_interpolated_by_cell.csv"))
```
  
```{r}
# And plot again
prices_interpolated_map_suitable <- price_interpolated_out %>%
  right_join(ocean_df, by = "cell_id") %>%
  dplyr::filter(!is.na(price)) %>%
  ggplot()+
  map_layer_moll+
  geom_raster(aes(x = lon, y = lat, fill = price))+
  scale_fill_viridis(name = "Price per Dive (USD)",
                      breaks = price_breaks,
                      limits = price_limits,
                     oob = squish,
                     option = "C",
                     guide = guide_colorbar(title.position = "top", title.hjust = 0.5, barwidth = barwidth, barheight = barheight))+
  theme_basic_map()+
  labs(x = "", y = "", title = paste0("Median dive price for all model pixels (USD): ",
                                      dollar(median(price_interpolated_out$price, na.rm = T))))
  
# Output
save_plots(prices_interpolated_map_suitable, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "price_interpolated_suitable_by_cell.png"),
           width_in = 7.5,
           height_in = 4.5)

prices_interpolated_map_suitable
```

And finally, let's combine our final four price plots into one to show the process of creating the sensitivity price layers. 

```{r}
top <- plot_grid(prices_country_region_map + theme(legend.position = "none"), 
                 prices_country_region_suitable_map + theme(legend.position = "none"), 
                 prices_interpolated_map_all + theme(legend.position = "none"), 
                 prices_interpolated_map_suitable + theme(legend.position = "none"),
                 ncol = 2,
                 nrow = 2,
                 labels = "AUTO",
                 label_y = 1.02)

prices_combined_map <- plot_grid(top,
                                 get_legend(prices_interpolated_map_suitable),
                                 ncol = 1,
                                 rel_heights = c(1,0.25))

# Output
save_plots(prices_combined_map, 
           file_name = file.path(emlab_project_dir, "figures", "model-inputs", "price_sensitivity_combo_plot.png"),
           width_in = 7.5,
           height_in = 6.5)

prices_combined_map
```

# Final Checks

Let's do a final check that we have the same number of cells with data in each of our model inputs. We can then filter these and save them as a .Rds object for Ren. 

```{r}
# Rename our 5 input layers
ocean_df_with_eezs <- cells_by_eez_out
dives_input_raw <- dives_by_cell_out
suitability_input_raw <- spots_by_cell
price_constant_input_raw <- price_constant
price_country_region_input_raw <- prices_by_cell_country_region_out
price_interpolated_input_raw <- price_interpolated_out

# Dives - 1814 rows
dives_input <- dives_input_raw %>%
  dplyr::filter(!is.na(n_dives_extrap) & ocean ==1) %>%
  dplyr::select(cell_id, n_dives_extrap, n_dives_extrap_min, n_dives_extrap_max)

nrow(dives_input)

# Suitability _ 1814 rows
suitability_input <- suitability_input_raw %>%
  dplyr::filter(suitable == 1) %>%
  dplyr::select(cell_id, suitable)

nrow(suitability_input)

# Prices constant  - 1814 rows
price_constant_input <- price_constant_input_raw %>%
  dplyr::filter(!is.na(price)) %>%
  dplyr::select(cell_id, price)

nrow(price_constant_input)

# Prices country/region - 1814 rows
price_country_region_input <- price_country_region_input_raw %>%
  dplyr::filter(!is.na(price)) %>%
  dplyr::select(cell_id, price, country_iso3 = eez_ter1_iso3, region)

nrow(price_country_region_input)

# Prices interpolated - 1814 rows
price_interpolated_input <- price_interpolated_input_raw %>%
  dplyr::filter(!is.na(price)) %>%
  dplyr::select(cell_id, price)

nrow(price_interpolated_input)

# And save
save(ocean_df_with_eezs,
     dives_input,
     suitability_input,
     price_constant_input,
     price_country_region_input,
     price_interpolated_input,
     file = file.path(emlab_project_dir, "data", "02-processed", "model-inputs", "tourism_model_input.RData"))
```

```{r}
# Which cells have dives, but no EEZ info? 425 cells...
cells_with_dives_but_no_eez_info <- dives_input %>%
  dplyr::filter(!is.na(n_dives_extrap)) %>%
  left_join(ocean_df_with_eezs, by = "cell_id") %>%
  dplyr::filter(is.na(mrgid))
```

