---
output:
  html_document: default

title: 'Price Data Collection Sheets - Week 1'
author: "Kat Millage"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
---

```{r setup, include=FALSE}
set.seed(123)

# Sheet Specific Packages
library(googlesheets4)

# Source common R file
source(here::here("common.R"))
```

# Introduction

This script generates the first set of price data collection sheets from the dive operator database.

```{r}
# Load data
diving_center_dat <- read_csv(file.path(emlab_project_dir, "data", "02-processed", "data-prep", "diving_center_database_clean.csv"))
```

## Diving Center Selection

We have 9909 diving centers in our database with URLs. For week 1, let's set the goal of collecting data on 5% of these shops. If this bleeds over into week 2 that's fine as well. 

```{r}
# Extract only shops with websites
diving_center_dat_good_url <- diving_center_dat %>%
  dplyr::filter(!is.na(website))

# How many shops would be 5%? 
n_5percent <- ceiling(0.05*nrow(diving_center_dat_good_url))

# Now let's randomly generate a testing sample. 
testing_sample <- diving_center_dat_good_url %>%
  dplyr::filter(!is.na(continent)) %>%
  group_by(continent) %>%
  sample_n(5, replace = T) %>%
  ungroup()

# Keep only the rows we need
testing_sample_keep <- testing_sample %>%
  dplyr::select(ID, business_name, website)
```

### Lesley

```{r}
# Now let's randomly generate a geographically representative sample for Lesley. Let's give her a sample comprising 2.75% of our total, and we'll make sure that they share 0.25% for QA/QC (works out to be 25 shops). 
n_2_75percent <- ceiling(0.0275*nrow(diving_center_dat_good_url))
n_0_25percent <- ceiling(0.0025*nrow(diving_center_dat_good_url))

# Extract sample 
lesley_sample <- diving_center_dat_good_url %>%
  dplyr::filter(!is.na(continent)) %>%
  group_by(continent) %>%
  sample_frac(0.0275, replace = T) %>%
  ungroup()

write_csv(lesley_sample, file.path(emlab_project_dir, "data", "01-raw", "prices", "lesley_sample_week1.csv"))

lesley_sample_keep <- lesley_sample %>%
  dplyr::select(ID, business_name, website)
```

### Bergen

Now for Bergen's sample, we want to give her 0.25% overlap with Lesley's sample so that I can check them. Let's do that first by randomly drawing 25 shops out of Lesley's sample. 

```{r}
# Randomly extract 0.25% overlap between the samples (actually about 10% for this sample)
bergen_qa_qc_sample <- lesley_sample %>%
  sample_n(n_0_25percent, replace = T)

write_csv(bergen_qa_qc_sample, file.path(emlab_project_dir, "data", "01-raw", "prices", "overlap_sample_week1.csv"))

# Size of database remaining
diving_center_dat_good_url_remaining <- diving_center_dat_good_url %>%
  dplyr::filter(!is.na(continent)) %>%
  anti_join(lesley_sample_keep %>% dplyr::select(ID), by = "ID") 

# Calculate the fraction we need to get the same sample size
frac_needed <- (n_2_75percent-nrow(bergen_qa_qc_sample))/nrow(diving_center_dat_good_url_remaining)

# Now let's extract the remaining sample for Bergen that's not in Lesley's sample and add the QA/QC rows
bergen_sample <- diving_center_dat_good_url_remaining %>%
  group_by(continent) %>%
  sample_frac(frac_needed, replace = T) %>%
  ungroup() %>%
  bind_rows(bergen_qa_qc_sample)

write_csv(bergen_sample, file.path(emlab_project_dir, "data", "01-raw", "prices", "bergen_sample_week1.csv"))

bergen_sample_keep <- bergen_sample %>%
  dplyr::select(ID, business_name, website)
```

## Make Google Sheets

### Lesley

```{r}
lesley_nrow <- nrow(lesley_sample_keep)

lesley_sheet_1_blank <- tibble(date_accessed = character(lesley_nrow),
                        original_site_language = character(lesley_nrow),
                        google_translate = character(lesley_nrow),
                        can_read = character(lesley_nrow),
                        business_name_manual = character(lesley_nrow),
                        country = character(lesley_nrow),
                        is_dive_related = character(lesley_nrow),
                        is_PADI = character(lesley_nrow),
                        is_NAUI = character(lesley_nrow),
                        is_SSI_TDI = character(lesley_nrow),
                        other_cert_agencies = character(lesley_nrow),
                        notes = character(lesley_nrow))

lesley_sheet_2_blank <- tibble(currency = character(lesley_nrow),
                        location_type = character(lesley_nrow),
                        p_type = character(lesley_nrow),
                        p_description = character(lesley_nrow),
                        p = character(lesley_nrow),
                        includes_equiptment = character(lesley_nrow),
                        includes_food_drink = character(lesley_nrow),
                        n_dives = character(lesley_nrow),
                        notes = character(lesley_nrow))

lesley_sheet_1 <- lesley_sample_keep %>%
  bind_cols(lesley_sheet_1_blank)

lesley_sheet_2 <- lesley_sample_keep %>%
  bind_cols(lesley_sheet_2_blank)
```

### Bergen

```{r}
bergen_nrow <- nrow(bergen_sample_keep)

bergen_sheet_1_blank <- tibble(date_accessed = character(bergen_nrow),
                        original_site_language = character(bergen_nrow),
                        google_translate = character(bergen_nrow),
                        can_read = character(bergen_nrow),
                        business_name_manual = character(bergen_nrow),
                        country = character(bergen_nrow),
                        is_dive_related = character(bergen_nrow),
                        is_PADI = character(bergen_nrow),
                        is_NAUI = character(bergen_nrow),
                        is_SSI_TDI = character(bergen_nrow),
                        other_cert_agencies = character(bergen_nrow),
                        notes = character(bergen_nrow))

bergen_sheet_2_blank <- tibble(currency = character(bergen_nrow),
                        location_type = character(bergen_nrow),
                        p_type = character(bergen_nrow),
                        p_description = character(bergen_nrow),
                        p = character(bergen_nrow),
                        includes_equiptment = character(bergen_nrow),
                        includes_food_drink = character(bergen_nrow),
                        n_dives = character(bergen_nrow),
                        notes = character(bergen_nrow))

bergen_sheet_1 <- bergen_sample_keep %>%
  bind_cols(bergen_sheet_1_blank)

bergen_sheet_2 <- bergen_sample_keep %>%
  bind_cols(bergen_sheet_2_blank)
```

### Send to Google Sheets

```{r}
# sheet_3 <- tibble(col1 = c("T", "F"),
#                   col2 = c("Certification Course", "Boat Dive", "Shore Dive", "Other"),
#                   col3 = c("Marine", "Freshwater", "Unknown"),
#                   col4 = c("Snorkel", "Discover Scuba", "Open Water", "Adv. Open Water", "Rescue", "1 Dive", "2 Dives", "N Dives"))
```


```{r}
lesley_data <- list(center_info = lesley_sheet_1, prices = lesley_sheet_2)
bergen_data <- list(center_info = bergen_sheet_1, prices = bergen_sheet_2)

lesley_gs <- gs4_create(
  "lesley_price_data_collection_week1",
  sheets = lesley_data)

bergen_gs <- gs4_create(
  "bergen_price_data_collection_week1",
  sheets = bergen_data)

```

