---
output:
  html_document: default

title: 'Dive Operator Data - Summary/Visualization'
author: "Kat Millage"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
---

```{r setup, include=FALSE}
set.seed(123)

# Script Specific Packages
library(treemap)

# Source common R file
source(here::here("common.R"))
```

# Introduction

This script summarizes and visualizes the general characteristics associated with the dive operator database, as well as statistics pertaining to those sampled as part of our price data collection.

```{r}
# Load cleaned dive center data (from Fabio et al.) and our cleaned shop metadata file
diving_center_dat<- read_csv(file.path(emlab_project_dir, "data", "02-processed", "data-prep", "diving_center_database_clean.csv"))

sampled_operator_dat <- read_csv(file.path(emlab_project_dir, "data", "02-processed", "data-prep", "sampled_operators_metadata_clean.csv"))
```

# Plot Settings & Setup

We're going to want to look at some of our data in a binned fashion - let's go ahead and bin our operators by location. 

```{r}
barwidth = 15
barheight = 1
```

```{r}
# Now let's extract what we need from our operator database
dive_operator_dat_spatial <- diving_center_dat %>%
  dplyr::select(lon = longitude, lat = latitude, ID)

# Make it a SPDF and transform
dive_operator_spdf <- SpatialPointsDataFrame(cbind(dive_operator_dat_spatial$lon, dive_operator_dat_spatial$lat), as.data.frame(dive_operator_dat_spatial$ID), proj4string=CRS("+proj=longlat"))

dive_operator_moll_df <- spTransform(dive_operator_spdf, CRS = prj_moll) %>%
  as.data.frame() %>%
  rename(ID = `dive_operator_dat_spatial.ID`,
         x_coord = `coords.x1`,
         y_coord = `coords.x2`)
```

```{r}
# Now let's set a bin size
bin_size <- 50000 # in meters
bin_size_txt <- "50km"

# Match Mollewiede coordinates back to the dive operator data and bin
dive_operator_dat_binned <- diving_center_dat %>%
  left_join(dive_operator_moll_df, by = "ID") %>%
  mutate(x_coord_cen = bin_spatial_data(x_coord, bin_size = bin_size),
         y_coord_cen = bin_spatial_data(y_coord, bin_size = bin_size))
```

# Summaries: Number of Dive Operators

We already made a few summary plots when we cleaned this database, but let's go ahead and make a few more pertaining to the total number of operators in the database provided to us by Schubauer et al. (in review). 

## By bin

Let's transform our lat/lon coordinates for our sampled dive operators to mollewide and then bin them into 50 x 50 km2 bins. 

```{r}
# Now we make a map of our number of dive operators
dive_operators_map <- dive_operator_dat_binned %>%
  group_by(x_coord_cen, y_coord_cen) %>%
  summarize(n_shops = n_distinct(ID)) %>%
  ggplot()+
  map_layer_moll +
  geom_raster(aes(x = x_coord_cen, y = y_coord_cen, fill = log10(n_shops)), show.legend = T)+
  scale_fill_viridis(option = "plasma", limits = c(0,2), oob = squish, labels = c("1", "25", "50", "75", "\u2265 100"))+
  theme(legend.position = "bottom",
        title = element_text(hjust = 0.5))+
  labs(title = paste0("Total number of dive operators: ", format(n_distinct(dive_operator_dat_binned$ID), big.mark = ",")),
    x = "", y = "")+
  guides(fill = guide_colorbar(barwidth = barwidth, barheight = barheight, title = "Dive operators (#)", title.position = "top", title.hjust = 0.5))+
  theme_basic()

# Output
save_plots(dive_operators_map, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_all_50km_50km_bins.png"),
           width_in = 7.5,
           height_in = 5)

dive_operators_map
```

## By country

```{r}
# Get our number of shops by country and reorder
dive_operators_country_data_all <- dive_operator_dat_binned %>%
  mutate(country_iso3 = case_when(country_scraped == "Saint Martin" ~ "MAF",
                                  T ~ countrycode(country_scraped, "country.name", "iso3c"))) %>%
  mutate(country_name = countrycode(country_iso3, "iso3c", "country.name"),
         continent = case_when(country_name == "Saint Martin (French part)" ~ "Americas",
                               T ~ continent)) %>%
  group_by(continent, country_name) %>%
  summarize(value = n_distinct(ID)) %>%
  ungroup() %>%
  arrange(continent, desc(value)) %>%
  rename(group = continent,
         country = country_name)

# Plot
dive_operators_country_plot_all <- dive_operators_country_data_all %>%
  ggplot() +
  geom_bar(aes(x=fct_reorder(country, value), y=value, fill=fct_rev(group)), stat="identity", alpha=1) +
  geom_text(aes(x=fct_reorder(country, value), y=value+10, label=country), color="black", fontface="bold",alpha=0.6, size=1.5, inherit.aes = FALSE, hjust = 0)+
  scale_y_continuous(expand = c(0,0), limits = c(0,ceiling(max(dive_operators_country_data_all$value))+100))+
  scale_fill_brewer(palette = "Set1", name = "", guide = guide_legend(reverse = T))+
  facet_grid(rows = "group", scales = "free_y", space = "free")+
  labs(y = "Dive operators (#)")+
  coord_flip()+
  theme_basic()+
  theme(
    legend.position = "none",
    axis.text.y  = element_blank(),
    axis.title.y = element_blank()
  )

# Output

save_plots(dive_operators_country_plot_all, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_all_by_country_and_continent.png"),
           width_in = 7.5,
           height_in = 9.5)

dive_operators_country_plot_all
```

## By country (top 25)

Now let's summarize the number of operators by country with a circle plot

```{r}
# Get our number of shops by country and reorder
dive_operators_country_data <- dive_operator_dat_binned %>%
  group_by(continent, sovereignt) %>%
  summarize(value = n_distinct(ID)) %>%
  ungroup() %>%
  arrange(continent, desc(value)) %>%
  mutate(group = fct_reorder(continent, value),
         country = fct_reorder2(sovereignt, continent, value)) %>%
  dplyr::select(-continent, -sovereignt) %>%
  dplyr::filter(value >= 25)

# Set a number of 'empty bars' to add at the end of each group
empty_bar <- 3
to_add <- data.frame( matrix(NA, empty_bar*nlevels(dive_operators_country_data$group), ncol(dive_operators_country_data)) )
colnames(to_add) <- colnames(dive_operators_country_data)
to_add$group <- rep(levels(dive_operators_country_data$group), each=empty_bar)
dive_operators_country_data <- rbind(dive_operators_country_data, to_add)
dive_operators_country_data <- dive_operators_country_data %>% arrange(group)
dive_operators_country_data$id <- seq(1, nrow(dive_operators_country_data))

# Get the name and the y position of each label
label_data <- dive_operators_country_data
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id-0.5) /number_of_bar
label_data$hjust <- ifelse( angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle+180, angle)

# prepare a data frame for base lines
base_data <- dive_operators_country_data %>% 
  group_by(group) %>% 
  summarize(start=min(id), end=max(id) - empty_bar) %>% 
  rowwise() %>% 
  mutate(title=mean(c(start, end)))
 
# prepare a data frame for grid (scales)
grid_data <- base_data
grid_data$end <- grid_data$end[c(nrow(grid_data), 1:nrow(grid_data)-1)] + 1
grid_data$start <- grid_data$start - 1
grid_data <- grid_data[-1,]

# Plot
dive_operators_country_plot <- dive_operators_country_data %>%
  ggplot(aes(x=as.factor(id), y=value, fill=fct_rev(group))) +

  geom_bar(aes(x=as.factor(id), y=value, fill=fct_rev(group)), stat="identity", alpha=0.5) +
  
  # Add lines. I do it at the beginning to make sure barplots are OVER them
  geom_segment(data=grid_data, aes(x = end, y = 800, xend = start, yend = 800), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 400, xend = start, yend = 400), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 200, xend = start, yend = 200), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 100, xend = start, yend = 100), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  geom_segment(data=grid_data, aes(x = end, y = 50, xend = start, yend = 50), colour = "grey", alpha=1, size=0.3 , inherit.aes = FALSE ) +
  
  # Add text showing the value of each 100/75/50/25 lines
  annotate("text", x = rep(max(dive_operators_country_data$id),5), y = c(50, 100, 200, 400, 800), label = c("50", "100", "200", "400", "800") , color="grey", size=1 , angle=0, fontface="bold", hjust=1) +
  
  geom_bar(aes(x=as.factor(id), y=value, fill=fct_rev(group)), stat="identity", alpha=1) +
  ylim(-100,1600) +
  theme_basic()+
  theme(
    legend.position = "none",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-3.75,4), "cm") 
  ) +
  coord_polar()+
  geom_text(data=label_data, aes(x=id, y=value+15, label=country, hjust=hjust), color="black", fontface="bold",alpha=0.6, size=1.5, angle= label_data$angle, inherit.aes = FALSE)
  #scale_fill_brewer(palette = "Set1", name = "")


# Output
save_plots(dive_operators_country_plot, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_all_by_country_and_continent_top_25_circle_plot.png"),
           width_in = 7.5,
           height_in = 6.5)

dive_operators_country_plot
```
# Summaries: Operators Meeting Criteria

Let's summarize the number of our sampled shops meeting the basic criteria for inclusion in our study (has a valid URL, is readable, is dive related, and offers marine dive services). We obviously don't know that these would be consistent across the non-sampled portion of the dataset, but it's a good place to start. 

## Total

```{r}
# Summarize our totals
 sampled_operator_validity <- sampled_operator_dat %>%
   count(is_valid_url,
         is_readable,
         is_dive_related,
         offers_marine_dive_services) %>%
  mutate(percent = round((n/n_distinct(sampled_operator_dat$ID))*100, 2))

 # Everything looks good!
 sampled_operator_validity_table <- basic_table_html(sampled_operator_validity %>% mutate_all(function(x)ifelse(is.na(x), "---", x)),
                                            column_names = c("Valid URL?", "Readable?", "Dive Related Business?", "Offers Marine Dive Services?", "Number", "Percent"),
                                            fig_caption = "")

 save_table(sampled_operator_validity_table,
            file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_sampled_validity_table.png"))
 
 sampled_operator_validity_table
```


```{r}
# Let's visualize this while we're here
sampled_operator_count <- sampled_operator_dat %>%
  mutate(group1 = case_when(is_valid_url == F | is_readable == F ~ "Invalid URL or incomprehensible",
                           is_dive_related == F ~ "Non dive-related businesses",
                           offers_marine_dive_services == F ~ "Does not offer marine dive services",
                           offers_marine_dive_services == T ~ "Offers marine dive services")) %>%
  group_by(group1) %>%
  summarize(count = n_distinct(ID))
  
# Now we plot it
png(filename=file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_sampled_validity_tree.png"),width=7,height=7,units = "in",res =300)
sampled_operator_validity_plot <- treemap::treemap(sampled_operator_count, #Your data frame object
        index=c("group1"),  #A list of your categorical variables
        vSize = "count",  #This is your quantitative variable
        type="index", #Type sets the organization and color scheme of your treemap
        palette = c("#e3e3aa", "#d69ca2", "#fabf9b", "#738deb"),
        title=paste0("Dive operators sampled: ", format(sum(sampled_operator_count$count), big.mark = ",")), #Customize your title
        fontsize.title = 16, #Change the font size of the title
        fontsize.labels=c(16,14),
        fontface.labels=c(2,2),
        #bg.labels=c("transparent"),
        align.labels=list(
        c("center", "center")),
        ymod.labels = list(0),
        inflate.labels = F,
        overlap.labels = 1,
        border.col=c("black"),
        border.lwds=c(4,2),
        fontcolor.labels = c("black"),
        position.legend = "none",
        mirror.x = T,
        mirror.y = T)
dev.off()
```

## By continent

Then by continent 

```{r}
# Now let's visualize
sampled_operator_count_continent <- sampled_operator_dat %>%
  mutate(group1 = case_when(is_valid_url == F | is_readable == F ~ "Invalid URL or incomprehensible",
                           is_dive_related == F ~ "Non dive-related businesses",
                           offers_marine_dive_services == F ~ "Does not offer marine dive services",
                           offers_marine_dive_services == T ~ "Offers marine dive services")) %>%
  group_by(group1, continent) %>%
  summarize(count = n_distinct(ID))

# Visualize
png(filename=file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_sampled_validity_tree_by_continent.png"),width=7,height=7,units = "in",res =300)
sampled_operator_validity_plot_continent <- treemap::treemap(sampled_operator_count_continent, #Your data frame object
        index=c("continent", "group1"),  #A list of your categorical variables
        vSize = "count",  #This is your quantitative variable
        type="categorical", #Type sets the organization and color scheme of your treemap
        vColor = "group1",
        palette = c("#e3e3aa", "#d69ca2", "#fabf9b", "#738deb"),
        title=paste0("Dive Operators Surveyed: ", format(sum(sampled_operator_count_continent$count), big.mark = ",")), #Customize your title
        fontsize.title = 16, #Change the font size of the title
        fontsize.labels=c(16,6.5),
        fontface.labels=c(2,2),
        #bg.labels=c("transparent"),
        align.labels=list(
        c("center", "center")),
        ymod.labels = list(0),
        inflate.labels = F,
        overlap.labels = 0,
        border.col=c("black", "black"),
        border.lwds=c(4,2),
        fontcolor.labels = c("black", "black"),
        position.legend = "none",
        mirror.x = T,
        mirror.y = T)
dev.off()
```

# Summaries: Sampling Rates

Now that we've got a good idea of our "false positive" rates across a few different metrics, let's look at our sampling rates spatially.

## By bin

Let's make a map of the % of operators in each bin sampled. 

```{r}
# Map of proportion of operators sampled by bin
dive_operators_prop_sampled_map <- dive_operator_dat_binned %>%
  mutate(was_sampled = ifelse(ID %in% sampled_operator_dat$ID,T, F)) %>%
  group_by(x_coord_cen, y_coord_cen) %>%
  summarize(n_shops = n_distinct(ID),
            n_shops_sampled = n_distinct(ID[was_sampled])) %>%
  ungroup() %>%
  mutate(prop_sampled = n_shops_sampled/n_shops) %>%
  ggplot()+
  map_layer_moll +
  geom_raster(aes(x = x_coord_cen, y = y_coord_cen, fill = prop_sampled), show.legend = T)+
  scale_fill_gradient(low = "lightblue4", high = "#f9b641", labels = percent)+
  theme(legend.position = "bottom",
        title = element_text(hjust = 0.5))+
  labs(x = "", y = "", 
       title = paste0("Total number of dive operators: ", format(n_distinct(dive_operator_dat_binned$ID), big.mark = ",")))+
  guides(fill = guide_colorbar(barwidth = barwidth, barheight = barheight, title = "Dive operators (% sampled)", title.position = "top", title.hjust = 0.5))+
  theme_basic()

# Output
save_plots(dive_operators_prop_sampled_map, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_sampled_50km_50km_bins.png"),
           width_in = 7.5,
           height_in = 5)

dive_operators_prop_sampled_map
```

## By business type

Then look at sampling rates by business type. 

```{r}
# Let's recreate one of our earlier figures looking at center types. This time we want to see the proportions of those we sampled. 
sampling_summary_business_type_abs <- diving_center_dat %>%
  mutate(was_sampled = case_when(ID %in% sampled_operator_dat$ID ~ T,
                                 T ~ F)) %>%
  group_by(type_of_business_new, was_sampled) %>%
  summarize(shops_n = n_distinct(ID)) %>%
  ungroup() %>%
  group_by(type_of_business_new) %>%
  mutate(shops_total = sum(shops_n, na.rm = T)) %>%
  ungroup() %>%
  mutate(label = ifelse(was_sampled, paste0(shops_n, " of ", shops_total, "\nsampled"), "")) %>%
  ggplot()+
  aes(x = reorder(type_of_business_new, shops_n), y = shops_n, fill = was_sampled)+
  geom_bar(stat = "identity")+
  geom_text(aes(label = label), hjust = -0.1, family = "Roboto", size = 2, color = "white")+
  scale_fill_manual(values = c("#042333", "#f9b641"))+
  coord_flip()+
  labs(x = "", y = "Dive operators (#)",
       title = paste0("Total number of dive operators: ", format(n_distinct(diving_center_dat$ID), big.mark = ",")))+
  guides(fill = guide_legend(title = "Sampled?", reverse = T))+
  theme_basic()

sampling_summary_business_type_abs
```

And then the same by proportion
  
```{r}  
# Let's get the relative proportions too
sampling_summary_business_type_prop <- diving_center_dat %>%
  mutate(was_sampled = case_when(ID %in% sampled_operator_dat$ID ~ T,
                                 T ~ F)) %>%
  group_by(type_of_business_new) %>%
  summarize(Sampled = n_distinct(ID[was_sampled])/n_distinct(ID),
            `Not Sampled` = n_distinct(ID[!was_sampled])/n_distinct(ID),
            shops_n = n_distinct(ID)) %>%
  ungroup() %>%
  gather(was_sampled, prop, 2:3) %>%
  mutate(was_sampled = case_when(was_sampled == "Sampled" ~ T,
                                 T ~ F)) %>%
  mutate(label = ifelse(was_sampled, paste0(round(prop, 4)*100, "%\nsampled"), "")) %>%
  ggplot()+
  aes(x = reorder(type_of_business_new, shops_n), y = prop, fill = was_sampled)+
  geom_bar(stat = "identity")+
  geom_text(aes(label = label), hjust = -0.2, family = "Roboto", size = 2, color = "white")+
  scale_fill_manual(values = c("#042333", "#f9b641"))+
  coord_flip()+
  labs(x = "", y = "Dive operators (% of total)")+
  guides(fill = guide_legend(title = "Sampled?", reverse = T))+
  scale_y_continuous(labels = scales::percent)+
  theme_basic()

sampling_summary_business_type_prop
```

```{r}
# And combine and save
sampling_summary_business_type_out <- plot_grid(sampling_summary_business_type_abs + theme(legend.position = "none"), 
                                                sampling_summary_business_type_prop, 
                                                labels = "AUTO", ncol = 1, align = "hv", rel_heights = c(1, 1))


save_plots(sampling_summary_business_type_out, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_sampled_by_type.png"),
           width_in = 7.5,
           height_in = 7)

sampling_summary_business_type_out
```

### By continent

```{r}
# Number of operators sampled by continent
sampling_summary_continent_abs <- diving_center_dat %>%
  mutate(was_sampled = case_when(ID %in% sampled_operator_dat$ID ~ T,
                                 T ~ F)) %>%
  group_by(continent, was_sampled) %>%
  summarize(shops_n = n_distinct(ID)) %>%
  ungroup() %>%
  group_by(continent) %>%
  mutate(shops_total = sum(shops_n, na.rm = T)) %>%
  ungroup() %>%
  mutate(label = ifelse(was_sampled, paste0(shops_n, " of ", shops_total, "\nsampled"), "")) %>%
  dplyr::filter(!is.na(continent)) %>%
  ggplot()+
  aes(x = reorder(continent, shops_total), y = shops_n, fill = was_sampled)+
  geom_bar(stat = "identity")+
  geom_text(aes(label = label), hjust = -0.1, family = "Roboto", size = 2, color = "white")+
  theme_bw()+
  coord_flip()+
  labs(x = "", y = "Dive operators (#)",
       title = paste0("Total number of dive operators: ", format(n_distinct(diving_center_dat$ID), big.mark = ",")))+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = c("#042333", "#f9b641"))+
  guides(fill = guide_legend(title = "Sampled?", reverse = T))+
  theme_basic()

sampling_summary_continent_abs
```
  
```{r}
# Let's get the relative proportions too
sampling_summary_continent_prop <- diving_center_dat %>%
  mutate(was_sampled = case_when(ID %in% sampled_operator_dat$ID ~ T,
                                 T ~ F)) %>%
  group_by(continent) %>%
  summarize(Sampled = n_distinct(ID[was_sampled])/n_distinct(ID),
            `Not Sampled` = n_distinct(ID[!was_sampled])/n_distinct(ID),
            shops_n = n_distinct(ID)) %>%
  ungroup() %>%
  dplyr::filter(!is.na(continent)) %>%
  gather(was_sampled, prop, 2:3) %>%
  mutate(was_sampled = case_when(was_sampled == "Sampled" ~ T,
                                 T ~ F)) %>%
  mutate(label = ifelse(was_sampled, paste0(round(prop, 4)*100, "%\nsampled"), "")) %>%
  ggplot()+
  aes(x = reorder(continent, shops_n), y = prop, fill = was_sampled)+
  geom_bar(stat = "identity")+
  geom_text(aes(label = label), hjust = -0.1, family = "Roboto", size = 2, color = "white")+
  theme_bw()+
  coord_flip()+
  labs(x = "", y = "Dive operators (% of total)")+
  scale_y_continuous(expand = c(0,0), labels = scales::percent)+
  scale_fill_manual(values = c("#042333", "#f9b641"))+
  guides(fill = guide_legend(title = "Sampled?", reverse = T))+
  theme_basic()

sampling_summary_continent_prop
```
```{r}
# And combine and save
sampling_summary_continent_out <- plot_grid(sampling_summary_continent_abs + theme(legend.position = "none"), 
                                            sampling_summary_continent_prop, 
                                            labels = "AUTO", ncol = 1, align = "hv", rel_heights = c(1, 1))

save_plots(sampling_summary_continent_out, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_sampled_by_continent.png"),
           width_in = 7.5,
           height_in = 7)

sampling_summary_continent_out
```

# Fig SX - Three panel plot summarizing price data collection

```{r}
fig_s1_b_legend <- get_legend(sampling_summary_continent_abs + theme(legend.box.margin = margin(12,0,0,0)))

fig_s1_b <- plot_grid(sampling_summary_continent_abs + theme(legend.position = "none") + labs(title = ""), 
                      sampling_summary_continent_prop + theme(legend.position = "none"), 
                      nrow = 1, align = "hv", rel_widths = c(1, 1), 
                      labels = c("B", "C"))

fig_s1 <- plot_grid(dive_operators_map,
                    fig_s1_b,
                    fig_s1_b_legend,
                    labels = c("A", "", ""),
                    nrow = 3, rel_heights = c(4,1.5,0.3), align = "hv")

save_plots(fig_s1, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "dive_operators_fig_s1.png"),
           width_in = 7.5,
           height_in = 7.5)
```

