---
output:
  html_document: default

title: 'Price Data - Summary/Visualization'
author: "Kat Millage"
date: "`r format(Sys.time(), '%B %d, %Y')`"
fontsize: 11pt
---

```{r setup, include=FALSE}
set.seed(123)

# Packages
library(ggridges)
library(gt)
library(gtsummary)

# Source common R file
source(here::here("common.R"))
```

# Introduction

This script summarizes and visualizes the price data collected from shops included in the dive operator database.

```{r}
# Load cleaned price dat
price_dat_day <- read_csv(file.path(emlab_project_dir, "data", "02-processed", "data-prep", "prices_day_trips_clean.csv"))
price_dat_liveaboard <- read_csv(file.path(emlab_project_dir, "data", "02-processed", "data-prep", "prices_liveaboards_clean.csv"))
```

# Plot Settings & Setup

```{r}
barwidth = 15
barheight = 1
```

```{r}
# Now let's extract what we need from our operator database
price_dat_day_spatial <- price_dat_day %>%
  dplyr::select(lon = longitude, lat = latitude, ID) %>%
  distinct()

# Make it a SPDF and transform
price_dat_spdf <- SpatialPointsDataFrame(cbind(price_dat_day_spatial$lon, price_dat_day_spatial$lat), as.data.frame(price_dat_day_spatial$ID), proj4string=CRS("+proj=longlat"))

price_dat_moll_df <- spTransform(price_dat_spdf, CRS = prj_moll) %>%
  as.data.frame() %>%
  rename(ID = `price_dat_day_spatial.ID`,
         x_coord = `coords.x1`,
         y_coord = `coords.x2`)
```

```{r}
# Now let's set a bin size
bin_size <- 50000 # in meters (50x50 km)
bin_size_txt <- "50km"

# Match Mollewiede coordinates back to the dive operator data and bin
price_dat_day_binned <- price_dat_day %>%
  left_join(price_dat_moll_df, by = "ID") %>%
  mutate(x_coord_cen = bin_spatial_data(x_coord, bin_size = bin_size),
         y_coord_cen = bin_spatial_data(y_coord, bin_size = bin_size))
```

# Summaries: Day Trips and Classes 

```{r}
# See what we've got for descriptions
unique(price_dat_day$description)
```

## Overall

Let's first make overall summaries for our single day dives and classes

```{r}
# Create a summary table by continent
price_dat_summary_table <- price_dat_day %>% 
  dplyr::filter(description %in% c("Snorkel", "1 Dive", "2 Dives", "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
   mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives", "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  dplyr::select(description, price_converted) %>%
  tbl_summary(., missing = "no",
              by = description,
              statistic = list(all_categorical() ~ "{n} / {N} ({p}%)"),
              label = list(price_converted ~ "Price (USD)")) %>%
  bold_labels()

# Save
price_dat_summary_table %>%
  as_gt() %>%
  gt::gtsave(filename = "prices_day_trips_and_courses.html", path = file.path(emlab_project_dir, "figures", "data-prep"))

price_dat_summary_table
```

```{r}
# Make a boxplot (most items)
price_dat_summary_ridges_all <- price_dat_day %>% 
  dplyr::filter(description %in% c("Snorkel", "1 Dive", "2 Dives", "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives", "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
 ggplot(aes(x = price_converted, y = description))+
  geom_density_ridges()+
  labs(y = "", x = "Price (USD)")+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_all, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_day_trips_and_courses_ridges.png"),
           width_in = 7.5,
           height_in = 4.5)

price_dat_summary_ridges_all
```

### Day trips only

```{r}
# Make a desnity plot (just dives)
price_dat_summary_ridges_dives_only <- price_dat_day %>% 
  dplyr::filter(description %in% c("1 Dive", "2 Dives", "Snorkel")) %>%
  mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives")) %>%
  ggplot(aes(x = price_converted, y = description))+
  geom_density_ridges()+
  labs(y = "", x = "Price (USD)")+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_dives_only, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_day_trips_ridges.png"),
           width_in = 7.5,
           height_in = 3)

price_dat_summary_ridges_dives_only
```

```{r}
# Density plot by boat/shore (just dives)
price_dat_summary_ridges_dives_only_type <- price_dat_day %>%
  dplyr::filter(description %in% c("1 Dive", "2 Dives", "Snorkel")) %>%
  mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives")) %>%
  ggplot(aes(x = price_converted, y = description))+
  geom_density_ridges(aes(fill = type), alpha = 0.7)+
  labs(y = "", x = "Price (USD)", fill = "Type")+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_dives_only_type, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_day_trips_by_type_ridges.png"),
           width_in = 7.5,
           height_in = 3.5)

price_dat_summary_ridges_dives_only_type
```

```{r}
# Density plot by equiptment (just dives)
price_dat_summary_ridges_dives_only_equipment <- price_dat_day %>%
  dplyr::filter(description %in% c("1 Dive", "2 Dives", "Snorkel")) %>%
  mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives")) %>%
  mutate(includes_full_equipment = fct_relevel(includes_full_equipment,
                                               "T", "F", "Unknown")) %>%
  ggplot(aes(x = price_converted, y = description))+
  geom_density_ridges(aes(fill = includes_full_equipment), alpha = 0.7)+
  labs(y = "", x = "Price (USD)", fill = "Includes all equipment?")+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_dives_only_equipment, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_day_trips_by_equipment_ridges.png"),
           width_in = 7.5,
           height_in = 3.5)

price_dat_summary_ridges_dives_only_equipment
```

### Courses only

```{r}
# Make a desnity plot (just courses)
price_dat_summary_ridges_courses_only <- price_dat_day %>% 
  dplyr::filter(description %in% c("Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  mutate(description = fct_relevel(description, 
                                   "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  ggplot(aes(x = price_converted, y = description))+
  geom_density_ridges()+
  labs(y = "", x = "Price (USD)")+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_courses_only, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_courses.png"),
           width_in = 7.5,
           height_in = 3)

price_dat_summary_ridges_courses_only
```

```{r}
# Density plot by boat/shore (just courses)
price_dat_summary_ridges_courses_only_type <- price_dat_day %>%
  dplyr::filter(description %in% c("Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  mutate(description = fct_relevel(description, 
                                   "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  ggplot(aes(x = price_converted, y = description))+
  geom_density_ridges(aes(fill = type), alpha = 0.7)+
  labs(y = "", x = "Price (USD)", fill = "Type")+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_courses_only_type, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_courses_by_type.png"),
           width_in = 7.5,
           height_in = 3.5)

price_dat_summary_ridges_courses_only_type
```

```{r}
# Density plot by equiptment (just dives)
price_dat_summary_ridges_courses_only_equipment <- price_dat_day %>%
  dplyr::filter(description %in% c("Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  mutate(description = fct_relevel(description, 
                                   "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  mutate(includes_full_equipment = fct_relevel(includes_full_equipment,
                                               "T", "F", "Unknown")) %>%
  ggplot(aes(x = price_converted, y = description))+
  geom_density_ridges(aes(fill = includes_full_equipment), alpha = 0.7)+
  labs(y = "", x = "Price (USD)", fill = "Includes all equipment?")+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_courses_only_equipment, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_courses_by_equipment_ridges.png"),
           width_in = 7.5,
           height_in = 3.5)

price_dat_summary_ridges_courses_only_equipment
```

## By Continent

```{r}
# Make a density plot (most items)
price_dat_summary_ridges_all_continent <- price_dat_day %>% 
  dplyr::filter(description %in% c("Snorkel", "1 Dive", "2 Dives", "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives", "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
 ggplot(aes(x = price_converted, y = description, fill = factor(continent, levels = c("Oceania", "Europe", "Asia", "Americas", "Africa"))))+
  geom_density_ridges(alpha = 0.7)+
  labs(y = "", x = "Price (USD)")+
  theme_minimal()+
  scale_fill_brewer(palette = "Set1", name = "", guide = guide_legend(reverse = T))+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_all_continent, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_day_trips_and_courses_by_continent.png"),
           width_in = 7.5,
           height_in = 5)

price_dat_summary_ridges_all_continent
```

### Day trips only

```{r}
# Make a desnity plot (just dives)
price_dat_summary_ridges_dives_only_continent <- price_dat_day %>% 
  dplyr::filter(description %in% c("1 Dive", "2 Dives", "Snorkel")) %>%
  mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives")) %>%
  ggplot(aes(x = price_converted, y = description, fill = factor(continent, levels = c("Oceania", "Europe", "Asia", "Americas", "Africa"))))+
  geom_density_ridges(alpha = 0.7)+
  labs(y = "", x = "Price (USD)")+
  scale_fill_brewer(palette = "Set1", name = "", guide = guide_legend(reverse = T))+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_dives_only_continent, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_day_trips_by_continent.png"),
           width_in = 7.5,
           height_in = 4.5)

price_dat_summary_ridges_dives_only_continent
```

```{r}
# Make a desnity plot (just courses)
price_dat_summary_ridges_courses_only_continent <- price_dat_day %>% 
  dplyr::filter(description %in% c("Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  mutate(description = fct_relevel(description, 
                                   "Discover Scuba (or Intro Dive)", "Open Water (or Level 1)", "Adv. Open Water (or Level 2)", "Rescue (or Level 3)")) %>%
  ggplot(aes(x = price_converted, y = description, fill = factor(continent, levels = c("Oceania", "Europe", "Asia", "Americas", "Africa"))))+
  geom_density_ridges(alpha = 0.7)+
  labs(y = "", x = "Price (USD)")+
  theme_minimal()+
  scale_fill_brewer(palette = "Set1", name = "", guide = guide_legend(reverse = T))+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_courses_only_continent, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_courses_by_continent.png"),
           width_in = 7.5,
           height_in = 4.5)

price_dat_summary_ridges_courses_only_continent
```

## By Country

```{r}
# Make a density plot (single dives only) by country
price_dat_summary_ridges_dives_only_country <- price_dat_day %>%
  dplyr::filter(description %in% c("1 Dive", "2 Dives", "Snorkel")) %>%
  mutate(description = fct_relevel(description, 
                                   "Snorkel", "1 Dive", "2 Dives")) %>%
  ggplot(aes(x = price_converted, y = country_name, fill = factor(continent, levels = c("Oceania", "Europe", "Asia", "Americas", "Africa"))))+
  geom_density_ridges(alpha = 0.8)+
  labs(y = "", x = "Price (USD)")+
  theme_minimal()+
  scale_fill_brewer(palette = "Set1", name = "", guide = guide_legend(reverse = T))+
  facet_wrap(~description)+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_dat_summary_ridges_dives_only_country, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_day_trips_by_country_and_continent.png"),
           width_in = 8.5,
           height_in = 8)

price_dat_summary_ridges_dives_only_country
```

## By Coordinates

Now let's look at it spatially

```{r}
# Calculate some binned stats
price_dat_binned_stats <- price_dat_day_binned %>%
  group_by(x_coord_cen, y_coord_cen) %>%
  summarize(med_price_snorkel = median(price_converted[description == "Snorkel"], na.rm = T),
            med_price_1_dive = median(price_converted[description == "1 Dive"], na.rm = T),
            med_price_2_dives = median(price_converted[description == "2 Dives"], na.rm = T)) %>%
  ungroup()
```

```{r}
# plot single dive prices
price_dat_binned_med_price_1_dive <- price_dat_binned_stats %>%
  dplyr::filter(!is.na(med_price_1_dive)) %>%
  ggplot()+
  map_layer_moll +
  geom_raster(aes(x = x_coord_cen, y = y_coord_cen, fill = med_price_1_dive), show.legend = T)+
  scale_fill_viridis(option = "plasma", breaks = c(0, 100, 200, 300), oob = squish, limits = c(0, 300), labels = dollar)+
  theme(legend.position = "bottom",
        title = element_text(hjust = 0.5))+
  guides(fill = guide_colorbar(barwidth = barwidth, barheight = barheight, title = "Median price - 1 tank dive (USD)", title.position = "top", title.hjust = 0.5))+
  theme_basic()+
  labs(x = "", y = "")

# Output
save_plots(price_dat_binned_med_price_1_dive, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_median_1_tank_50km_50km_bins.png"),
           width_in = 7.5,
           height_in = 5)

price_dat_binned_med_price_1_dive
```

```{r}
# plot 2 dive prices
price_dat_binned_med_price_2_dive <- price_dat_binned_stats %>%
  dplyr::filter(!is.na(med_price_2_dives)) %>%
  ggplot()+
  map_layer_moll +
  geom_raster(aes(x = x_coord_cen, y = y_coord_cen, fill = med_price_1_dive), show.legend = T)+
  scale_fill_viridis(option = "plasma", breaks = c(0, 100, 200, 300), oob = squish, limits = c(0, 300), labels = dollar)+
  theme(legend.position = "bottom",
        title = element_text(hjust = 0.5))+
  guides(fill = guide_colorbar(barwidth = barwidth, barheight = barheight, title = "Median price - 2 tank dive (USD)", title.position = "top", title.hjust = 0.5))+
  theme_basic()+
  labs(x = "", y = "")

# Output
save_plots(price_dat_binned_med_price_2_dive, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_median_2_tank_50km_50km_bins.png"),
           width_in = 7.5,
           height_in = 5)

price_dat_binned_med_price_2_dive
```

# Summaries: Price per Dive

For our input raster, we're going to combine prices for 1-tank, 2-tank, and N-tank dives to get the "per dive" price so we have greater spatial coverage of our price dataset. It turns out that these prices are actually significantly different, so we'll only keep the "best" per dive price for each operator, giving first priority to the price for a 1-tank dive if it's available, followed by the 2-tank dive price, and then the N-tank dive price. 

```{r}
# Let's first get the entries where we have 1-tank data - 262 entries
price_dat_1_dive <- price_dat_day %>%
  dplyr::filter(description %in% c("1 Dive")) %>%
  mutate(per_dive_price_best = price_converted)

# Then 2-tank data for operators where we don't have 1-tank data - 102 entries
price_dat_2_dive <- price_dat_day %>%
  dplyr::filter(description %in% c("2 Dives")) %>%
  dplyr::filter(!(ID %in% price_dat_1_dive$ID)) %>%
  mutate(per_dive_price_best = price_converted/2)

# Then N-tank data for operators where we don't have 1-tank or 2-tank data - 6 entries
price_dat_n_dive <- price_dat_day %>%
  dplyr::filter(description %in% c("N Dives")) %>%
  dplyr::filter(!(ID %in% price_dat_1_dive$ID) & !(ID %in% price_dat_2_dive$ID)) %>%
  mutate(per_dive_price_best = price_converted/n_dives) %>%
  dplyr::filter(!is.na(per_dive_price_best))

# Combine
price_dat_per_dive <- price_dat_1_dive %>%
  bind_rows(price_dat_2_dive) %>%
  bind_rows(price_dat_n_dive)
``` 

```{r}
# Including N-tank dive prices
anova <- aov(per_dive_price_best ~ description, data = price_dat_per_dive)
summary(anova)

# Only 1 and 2 tank dives
price_dat_per_dive_no_n <- price_dat_per_dive %>%
  dplyr::filter(description != "N Dives")

t_test <- t.test(price_dat_per_dive_no_n$per_dive_price_best ~ price_dat_per_dive_no_n$description)
t_test
```

```{r}
# Get quantiles 
price_summary_quantiles <- quantile(price_dat_per_dive$per_dive_price_best,
         probs = c(0, 0.01, 0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95, 0.99, 1))

price_summary_quantiles_table <- tibble() %>%
  bind_rows(price_summary_quantiles) %>%
  gather("quantile", "price", 1:11) %>%
  mutate(price = round(price, 2))

price_quantiles_table <- basic_table_html(price_summary_quantiles_table,
                                          column_names = c("Quantile", "Price Per Dive (USD)"),
                                          fig_caption = "")

save_table(price_quantiles_table,
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_per_dive_quantiles_table.png"))
 
price_quantiles_table
```

## By Country

```{r}
# Make a density plot of prices per dive by country
price_per_dive_country_plot <- price_dat_per_dive %>%
  ggplot(aes(x = per_dive_price_best, y = country_name, fill = factor(continent, levels = c("Oceania", "Europe", "Asia", "Americas", "Africa"))))+
  geom_density_ridges(alpha = 0.8)+
  labs(y = "", x = "Price per Dive (USD)")+
  scale_fill_brewer(palette = "Set1", name = "", guide = guide_legend(reverse = T))+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_per_dive_country_plot, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_per_dive_by_country_and_continent.png"),
           width_in = 8.5,
           height_in = 8)

price_per_dive_country_plot
```

```{r}
# Make a density plot of prices per dive by country
price_per_dive_country_plot_bar <- price_dat_per_dive %>%
  ggplot(aes(x = per_dive_price_best, y = country_name, fill = factor(continent, levels = c("Oceania", "Europe", "Asia", "Americas", "Africa"))))+
  geom_boxplot()+
  labs(y = "", x = "Price per Dive (USD)")+
  scale_fill_brewer(palette = "Set1", name = "", guide = guide_legend(reverse = T))+
  scale_x_continuous(label = scales::dollar)+
  theme_basic()

# Output
save_plots(price_per_dive_country_plot_bar, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_per_dive_by_country_and_continent_boxplot.png"),
           width_in = 8.5,
           height_in = 8)

price_per_dive_country_plot_bar
```

# Summaries: Liveaboards 

## By Destination (Total Price)

```{r}
# See what we've got for descriptions
unique(price_dat_liveaboard$dive_location)

# Do some wrangling
price_dat_liveaboard_edited <- price_dat_liveaboard %>%
  mutate(n_dives = case_when(!is.na(n_dives_total) ~ n_dives_total,
                             n_dives_per_day != "unlimited" ~ n_dive_days * as.numeric(n_dives_per_day)),
         min_price_per_dive = min_price/n_dives,
         max_price_per_dive = max_price/n_dives) 

# Get prices by location
price_dat_liveaboard_location <- price_dat_liveaboard_edited %>%
  group_by(dive_location) %>%
  summarize(min_price = min(min_price, na.rm = T),
            median_price = median(c(min_price, max_price), na.rm = T),
         max_price = max(max_price, na.rm = T),
         min_price_per_dive = min(min_price_per_dive, na.rm = T),
         median_price_per_dive = median(c(min_price_per_dive, max_price_per_dive), na.rm = T),
         max_price_per_dive = max(max_price_per_dive, na.rm = T)) %>%
  dplyr::filter(min_price_per_dive != Inf) %>%
  dplyr::filter(dive_location != "Guadalupe Island")

# And get the totals
price_dat_liveaboard_total <- price_dat_liveaboard_edited %>%
  summarize(min_price = min(min_price, na.rm = T),
            median_price = median(c(min_price, max_price), na.rm = T),
            max_price = max(max_price, na.rm = T),
            min_price_per_dive = min(min_price_per_dive, na.rm = T),
            median_price_per_dive = median(c(min_price_per_dive, max_price_per_dive), na.rm = T),
            max_price_per_dive = max(max_price_per_dive, na.rm = T)) %>%
  dplyr::filter(min_price_per_dive != Inf)%>%
  mutate(dive_location = "All")

# Join back together
price_dat_liveaboard_location <- price_dat_liveaboard_location %>%
  bind_rows(price_dat_liveaboard_total) 
```

```{r}
price_range_liveaboard_location_total <- price_dat_liveaboard_location %>%
  ggplot(aes(x = dive_location, y = median_price))+
  geom_crossbar(aes(ymin = min_price, ymax = max_price), width = 0.5, fill = "red", color = "black")+
  coord_flip()+
  labs(x = "Location", y = "Total Price (USD)")+
  scale_y_continuous(labels = dollar)+
  theme_basic()

# Output
save_plots(price_range_liveaboard_location_total, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_liveaboard_totals_by_location.png"),
           width_in = 8.5,
           height_in = 8)

price_range_liveaboard_location_total
```

```{r}
price_range_liveaboard_location_per_dive <- price_dat_liveaboard_location %>%
  ggplot(aes(x = dive_location, y = median_price_per_dive))+
  geom_crossbar(aes(ymin = min_price_per_dive, ymax = max_price_per_dive), width = 0.5, fill = "red", color = "black")+
  coord_flip()+
  labs(x = "Location", y = "Price per Dive (USD)")+
  scale_y_continuous(labels = dollar)+
  theme_basic()

# Output
save_plots(price_range_liveaboard_location_per_dive, 
           file_name = file.path(emlab_project_dir, "figures", "data-prep", "prices_liveaboard_per_dive_by_location.png"),
           width_in = 8.5,
           height_in = 8)

price_range_liveaboard_location_per_dive
```

